---
title: "COVID-19 Deaths Model"
output:
  html_document:
    df_print: paged
---

## Load theLibraries
```{r}
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(readr)
library(zoo)
library(scales)
library(sf)
library(urbnmapr)
library(plotly)
library(tidyverse)
library(mlbench)
library(caret)
```

## Merge census data with JHU data
```{r}
usfacts_data <- read.csv("census_data_0718.csv")
cases <- counties_sf
pop_density <- read.csv("USA_Population_Density.csv")
names(cases)[7] <- "county_fips_code"
cases_new <- merge(usfacts_data, cases, by="county_fips_code")
dim(cases_new)
cases_new <- merge(cases_new, pop_density, by="county_fips_code")
#Make characters factor for analysis
cases_new <- cases_new %>% mutate_if(is.character, factor)
dim(cases_new)
```

## Select Variables
```{r}
deaths_sel <- cases_new %>% select(county_name.x, state, total_pop, pop_density,        
  male_pop, female_pop, median_age, white_pop, 
  black_pop, asian_pop, hispanic_pop, amerindian_pop,
  commuters_by_public_transportation, median_income, 
  employed_pop, unemployed_pop, deaths_category)

summary(deaths_sel)

```

## Normalize by population
```{r}
deaths_sel <- deaths_sel %>% mutate(
  female_pop = female_pop / total_pop,
  male_pop = male_pop / total_pop,
  white_pop = white_pop / total_pop, 
  black_pop = black_pop / total_pop, 
  asian_pop = asian_pop / total_pop, 
  hispanic_pop = hispanic_pop / total_pop, 
  amerindian_pop = amerindian_pop / total_pop,
  commuters_by_public_transportation = commuters_by_public_transportation/ total_pop, 
  employed_pop = employed_pop / total_pop, 
  unemployed_pop = unemployed_pop / total_pop, 
  )

summary(deaths_sel)
```
```{r}
table(complete.cases(deaths_sel))
```


## Focus on states with Covid-19 outbreaks
```{r}
deaths_sel %>% pull(deaths_category) %>% table()
```


## Filter Hard Hit States - Cases
```{r}
deaths_sel %>% group_by(state) %>% 
  summarize(high_pct = sum(deaths_category == "High")/n()) %>%
  arrange(desc(high_pct))
```

## Designate States to use for Training model
```{r}
deaths_train <- deaths_sel %>% filter(
  state == "NY" |
    state == "CA" |
    state == "FL")

deaths_train %>% pull(deaths_category) %>% table()

```

## Plot Map
```{r}
counties_deaths <- as_tibble(map_data("county"))
counties_deaths <- counties_deaths %>% 
  rename(c(county = subregion, state = region)) %>%
  mutate(state = state.abb[match(state, tolower(state.name))]) %>%
  select(state, county, long, lat, group)
counties_deaths
```

## Add Variables to Map Data
```{r}
deaths_all <- counties_deaths %>% left_join(deaths_train %>% 
    mutate(county = county_name.x %>% str_to_lower() %>% 
        str_replace('\\s+county\\s*$', '')))
```

```{r}
ggplot(deaths_all, aes(long, lat)) + 
  geom_polygon(aes(group = group, fill = deaths_category)) + 
    coord_quickmap() 
```

## Check Variable Importance
```{r}
library(FSelector)
deaths_weight <- deaths_train %>% select(-county_name.x, -state, - total_pop)
deaths_weight <- deaths_weight %>% chi.squared(deaths_category ~ ., data = .) %>%
  as_tibble(rownames = "feature") %>%
  arrange(desc(attr_importance))
deaths_weight


```
```{r}
ggplot(deaths_weight,
  aes(x = attr_importance, y = reorder(feature, attr_importance))) +
  geom_bar(stat = "identity") +
  xlab("Importance score") + ylab("Feature")
```


## Get the 5 best features: 
```{r}
subset_tx <- cutoff.k(deaths_weight %>% column_to_rownames("feature"), 5)
subset_tx
```
## Univariate importance scores

```{r}
library(rpart)
library(rpart.plot)
#deaths_tx_new <- deaths_tx %>% select(-county_name.x, - total_pop, -state)
deaths_train %>% gain.ratio(deaths_category ~ ., data = .) %>%
  as_tibble(rownames = "feature") %>%
  arrange(desc(attr_importance))
```


## Build a Model - Decision Tree 
```{r}
library(caret)
fit_deaths <- deaths_train %>%
  train(deaths_category ~. - county_name.x - state - total_pop,
    data = . ,
    method = "rpart",
    #method = "rf",
    #method = "nb",
    trControl = trainControl(method = "cv", number = 10),
    tuneLength = 5)
fit_deaths
```


```{r}
library(rpart.plot)
rpart.plot(fit_deaths$finalModel, extra = 2)
varImp(fit_deaths)
```

## Build a Model - C.45
```{r}
C45fit_deaths <- deaths_train %>%
train(deaths_category ~ . - county_name.x - state - total_pop,
    data = . ,
    method = "J48",
    trControl = trainControl(method = "cv", number = 10)
    )
C45fit_deaths
```


```{r}
C45fit_deaths$finalModel
```

## Build a Model - Conditional Inference Tree
```{r}
ctreeFit_deaths <- deaths_train %>%
train(deaths_category ~ . - county_name.x - state - total_pop, 
    data = . ,
    method = "ctree",
    trControl = trainControl(method = "cv", number = 10)
    )
ctreeFit_deaths
```


```{r}
plot(ctreeFit_deaths$finalModel)
```

```{r}
predict(ctreeFit_deaths, head(cases_train))
```

## Build a Model - PART
```{r}
#Remove county name since it has too many levels which will make the code really slow
rulesfit_deaths <- deaths_train %>%
train(deaths_category ~ . - county_name.x - state - total_pop,
    data = . ,
    method = "PART",
    trControl = trainControl(method = "cv", number = 10)
    )
rulesfit_deaths
```

```{r}
rulesfit_deaths$finalModel
```

## Build a Model - Random Forest Fit
```{r}
randomForestFit_deaths <- deaths_train %>% 
  train(deaths_category ~ .  - county_name.x - state - total_pop,
    data = . ,
    method = "rf",
    trControl = trainControl(method = "cv", number = 10)
    )
randomForestFit_deaths
```


```{r}
randomForestFit_deaths$finalModel
```


## Artificial Neural Network
```{r}
nnetFit_deaths <- deaths_train %>% train(deaths_category ~ . - county_name.x - state - total_pop,
  method = "nnet",
  data = .,
    tuneLength = 5,
    trControl = trainControl(method = "cv", number = 10),
  trace = FALSE)
nnetFit_deaths
```

## Compare Models
```{r}
resamps <- resamples(list(
  decision = fit_deaths ,
  ctree = ctreeFit_deaths,
  rules = rulesfit_deaths,
  randomForest = randomForestFit_deaths,
  NeuralNet = nnetFit_deaths, 
  C45fit = C45fit_deaths
    ))
resamps
summary(resamps)
```

## Use Random Forest Classification Model to Predict Deaths Risk for all of US
```{r}
deaths_pred <- na.omit(deaths_sel)
deaths_pred$deaths_category <- predict(randomForestFit_deaths, deaths_pred)
deaths_pred <- counties_deaths %>% left_join(deaths_pred %>% 
    mutate(county = county_name.x %>% str_to_lower() %>% 
        str_replace('\\s+county\\s*$', '')))
```


## Plot the Prediction Results for US
```{r}
ggplot(deaths_pred, aes(long, lat)) + 
  geom_polygon(aes(group = group, fill = deaths_category)) + 
    coord_quickmap()
```

## Generate csv file for Prediction Results
```{r}
write.csv(deaths_pred, "C:/Users/Starr/Google Drive (starr.corbin@gmail.com)/R/covid19_prediction/prediction_results/deaths_prediction_results.csv")
```
