---
title: "Peak Cases and Deaths Classification Mapping"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Starr Corbin"
date: "7/24/2020"
output: html_document
---

##Load the R Libraries
```{r}
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(readr)
library(scales)
library(sf)
library(urbnmapr)
library(plotly)
library(tidyverse)
library(mlbench)
library(caret)
library(zoo)

```

##Pull the latest COVID-19 Data and Census Data
```{r}
jhu_url <- paste("https://raw.githubusercontent.com/CSSEGISandData/",
                 "COVID-19/master/csse_covid_19_data/", "csse_covid_19_time_series/",
                 "time_series_covid19_confirmed_US.csv", sep = "")

jhu_deaths <- paste("https://raw.githubusercontent.com/CSSEGISandData/",
                 "COVID-19/master/csse_covid_19_data/", "csse_covid_19_time_series/",
                 "time_series_covid19_deaths_US.csv", sep = "")

census_url <- paste("https://www2.census.gov/programs-surveys/popest/",
                    "datasets/2010-2019/counties/totals/co-est2019-alldata.csv",sep = "")
```

##Read in the COVID Data
```{r}
covidData <-
  read_csv(jhu_url) %>%
  rename(province = "Province_State",
         country_region = "Country_Region",
         county="Admin2")  %>%
  select(-c(UID,iso2,iso3,code3,country_region,Lat,Long_,Combined_Key)) %>% 
  pivot_longer(-c(province,county,FIPS), names_to = "d", 
               values_to = "cumulative_cases") %>%
  separate(d,c("Month","Day","Year"),sep="/") %>%
  mutate(dstring=sprintf("%02i/%02i/%02i",   #some parsing to make dates work correctly
                         as.numeric(Month), 
                         as.numeric(Day), 
                         as.numeric(Year)),
         d=as.Date(dstring,"%m/%d/%y")) %>%
  select(d,county,province,FIPS,cumulative_cases) %>%
  arrange(d) %>%
  group_by(FIPS) %>%  #for each FIPS ID, calculate new cases per day
  mutate(new_cases = cumulative_cases-dplyr::lag(cumulative_cases,1)) %>%
  ungroup() %>%
  filter(!is.na(FIPS))
```
```{r}
#Read in the COVID death data
covidDeath <-
  read_csv(jhu_deaths) %>%
  rename(province = "Province_State",
         country_region = "Country_Region",
         county="Admin2")  %>%
  select(-c(UID,iso2,iso3,code3,country_region,Lat,Long_,Combined_Key)) %>% 
  pivot_longer(-c(province,Population,county,FIPS), names_to = "d", 
               values_to = "cumulative_deaths") %>%
  separate(d,c("Month","Day","Year"),sep="/") %>%
  mutate(dstring=sprintf("%02i/%02i/%02i",   #some parsing to make dates work correctly
                         as.numeric(Month), 
                         as.numeric(Day), 
                         as.numeric(Year)),
         d=as.Date(dstring,"%m/%d/%y")) %>%
  select(d,county,province,FIPS,cumulative_deaths) %>%
  arrange(d) %>%
  group_by(FIPS) %>%  #for each FIPS ID, calculate new cases per day
  mutate(new_deaths = cumulative_deaths-dplyr::lag(cumulative_deaths,1)) %>%
  ungroup() %>%
  filter(!is.na(FIPS))
```
```{r}
covidData <- 
  covidData %>%
  left_join(select(covidDeath,d,FIPS,cumulative_deaths,new_deaths),
            by = c("FIPS","d"),
            na_matches="never")

rm(covidDeath)
```

##Read in the Population Data and Combine with COVID-19 Data
```{r}
co_est2019_alldata =
  read_csv(census_url) %>%
  select(STATE,COUNTY,POPESTIMATE2019) %>%
  rename(population="POPESTIMATE2019") %>%
  mutate(FIPS=as.numeric(str_c(as.character(STATE),as.character(COUNTY)))) 

#Combine covid and population data
covidData =
  covidData %>%
  left_join(select(co_est2019_alldata,FIPS,population),by="FIPS") %>%
  filter(!is.na(population))
```

##Calculate 7 Day Rolling Averages
```{r}
dCovid = covidData %>%
  mutate(ncap=new_cases/population*1e5) %>%
  mutate(dcap=new_deaths/population*1e5) %>%
  group_by(FIPS) %>%
  mutate(rnew=rollmean(new_cases,7,fill=NA,align="right")) %>%
  mutate(rdnew=rollmean(new_deaths,7,fill=NA,align="right")) %>%
  mutate(rncap=rollmean(ncap,7,fill=NA,align="right")) %>%
  mutate(rdcap=rollmean(dcap,7,fill=NA,align="right")) %>%
  ungroup()
```
```{r}
summary(dCovid)
```

#Extract the peak value occurred in the rolling average of new cases
```{r}
dCovid =
  dCovid %>%
  group_by(FIPS) %>%
  mutate(peak_rnew = max(rnew,na.rm=TRUE),
         peak_rdnew = max(rdnew,na.rm=TRUE),
         dateMax =d[which.max(rnew)],
         dateMax_rdnew =d[which.max(rdnew)],
         daySince = as.numeric(d - dateMax),
         daySince_rdnew = as.numeric(d - dateMax_rdnew))
```

##Take data corresponding to the latest date in the COIVD database.
```{r}
dNow = max(dCovid$d)
ddCovid=dCovid %>%
  filter(d==dNow)
```

##Read in and parse the county shapefile.
```{r}
counties_sf =
  get_urbn_map("counties", sf = TRUE) %>%
  mutate(FIPS=as.numeric(county_fips)) %>%
  left_join(ddCovid,by="FIPS")

```

###Combine county shape files into state features for plotting
```{r}
states <-
  counties_sf %>%
  group_by(province) %>%
  summarise(do_union = TRUE) %>%
  st_cast()
```

##Create the plot: Time since X
```{r}
pday_cases=counties_sf %>%
  ggplot() +
  geom_sf(aes(fill=daySince),size=0.1) +
  geom_sf(data=states,fill=NA,size=0.5) +
  scale_fill_gradient2(low="red",mid="yellow",high="green",midpoint=90/2,
                       breaks=c(0,30,60,90),
                       labels=c("0","30","60","90+"),
                       limits=c(0,90),
                       oob=squish)+
  ggtitle(str_c("Time Since Peak New Cases | ",dNow)) +
  coord_sf(crs=5070) +
  theme_bw() +
  labs(fill = "Days")

print(pday_cases) 
```
```{r}
#Create the plot
pday_deaths=counties_sf %>%
  ggplot() +
  geom_sf(aes(fill=daySince_rdnew),size=0.1) +
  geom_sf(data=states,fill=NA,size=0.5) +
  scale_fill_gradient2(low="red",mid="yellow",high="green",midpoint=90/2,
                       breaks=c(0,30,60,90),
                       labels=c("0","30","60","90+"),
                       limits=c(0,90),
                       oob=squish)+
  ggtitle(str_c("Time Since Peak New Deaths | ",dNow)) +
  coord_sf(crs=5070) +
  theme_bw() +
  labs(fill = "Days")

print(pday_deaths)
```

#Categorize days since last case into Low, Med, or High
```{r}
counties_sf$cases_category <- cut(counties_sf$daySince, breaks=c(-Inf, 30, 60, Inf), labels = c("High", "Med", "Low"))
```

##Categorize days since last death into Low, Med, or High
```{r}
counties_sf$deaths_category <- cut(counties_sf$daySince_rdnew, breaks=c(-Inf, 30, 60, Inf), labels = c("High", "Med", "Low"))
deaths_category <- cut(counties_sf$daySince_rdnew, breaks=c(-Inf, 30, 60, Inf), labels = c("High", "Med", "Low"))
```



